### 2019-03-06 paper.tex
Added some definitions worked on lemma.

\documentclass[12pt]{article}

%% bibliography stuff -- this needs come before the preamble inclusion
\usepackage[backend=bibtex,sorting=none]{biblatex}
\bibliography{\string~/Documents/academics/global_academics/global_bib.bib}
\usepackage{hyperref}

\def\gc{\overset{\text{GC}}{\rightarrow}}  % Granger causality arrow
\def\pwgc{\overset{\text{PW}}{\rightarrow}}  % Granger causality arrow
\def\te{\overset{\mathcal{T}}{\rightarrow}}  % Transfer entropy arrow
\def\gcg{\mathcal{G}}  % Granger-causality graph
\def\VAR{\mathsf{VAR}}  % VAR(p) model

\newcommand{\linE}[2]{\hat{\E}[#1\ |\ #2]}  % Linear projection
\newcommand{\linEerr}[2]{\xi[#1\ |\ #2]}  % Error of linear projection
\newcommand{\pa}[1]{P(#1)}  % Parents of a node
\newcommand{\anc}[1]{\mathcal{A}(#1)}  % Ancestors of a node

\input{\string~/Documents/academics/global_academics/latex_preamble}

\graphicspath{{./figures/}}

\title{Structure Learning for $VAR(p)$ Models}
\author{R. J. Kinnear, R. R. Mazumdar}

\begin{document}
\maketitle
\abstract{We study Granger Causality and propose a structure learning
  heuristic for uncovering a parsimonious representation of large
  $\mathsf{VAR}(p)$ models.}

\section{Introduction and Review}
\label{sec:introduction}
Consider a collection of stochastic processes producing observations
at discrete time intervals.  Are the underlying processes dependent?
Can we quantify any of the underlying relationships?  Can the arrow of
time help us to distinguish a directionality or flow of dependence
among our observed series?  In this paper we contribute to the
understanding of the notion of Granger-Causality
\cite{granger1969investigating} as a tool for answering these questions.

Though the notion of causality is a philosophically slippery concept,
it is fundamental to the way we understand the world and to the
progress of science in general.  Indeed, without faith in the
consistency of causal interactions the results of experimental science
could not be generalized or applied in any meaningful way.  In the
case of Granger-Causality, we state that if an event '$A$' provides us
with unique (that is, not available anywhere else) information about a
later event $B$, then $A$ must have a causal impact on $B$.  As
opposed to the notion of Causation promoted by Pearl
\cite{pearl2000art}, this is an entirely model-free notion of cause,
and instead leverages the intuition that a cause must precede it's
effect.

In practice, Granger's notion of causation it not a convincing test
for \textit{true} causation, since our statements about causation are
highly dependent upon the data that we are able to observe.  We prefer
instead to interpret Granger causality as a means of uncovering a flow
of ``information'' or ``energy'' through some underlying graph of
interactions.  Though this graph cannot be observed directly, we will
infer it's presence as a latent structure among our observed time
series data.

Finding the ``best'' graph structure consistent with observed data is
generally an extremely challenging problem, though the comparison of
quality between different structures, and hence the notion of
``best'', needs to be quantified.  In applications where we are
interested merely in minimizing the mean squared error of a linear
one-step-ahead predictor, then we will naturally desire an entirely
dense graph of connections, since each edge can only serve to reduce
estimation error.  However, since the number of edges scales
quadratically in $n$, it is imperative to infer a sparse causality
graph for large systems, both to avoid overfitting observed data, as
well as to aid the interpretability of the results.

In \cite{bach2004learning} the authors apply a local search heuristic
to an AIC penalized approximation of the likelihood where at each
iteration an edge is either added, removed, or reversed.  This is a
common approach to combinatorial optimization due to it's simplicity,
but is liable to get stuck in shallow local minima.

A second and wildly successful heuristic is the LASSO regularizer
\cite{tibshirani1996regression}, which can be understood as a natural
convex relaxation to penalizing the count of the non-zero edges.  The
LASSO enjoys strong theoretical guarantees \cite{wainwright2009sharp},
even in the case of time series data \cite{basu2015}
\cite{wong2016lasso}.

From a Bayesian perspective, spike-and-slab priors can be used to
estimate the probability of an edge's inclusion, after which a graph
can be chosen heuristically by thresholding or searching for the MAP
graph structure.

Our contribution is to provide an alternative heuristic based on
minimum spanning trees, inspired by our result in section
\ref{sec:theory} that the true causality graph can be recovered from
pairwise information alone in the case that the underlying graph is a
directed forest.  We compare this algorithm to a grouped LASSO
\cite{yuan2006model}, local search, and a Bayesian posterior
probability thresholding scheme.

%% -Can we also formulate an SDP relaxation?
%% -Fit trees jointly or via coordinate descent?

\section{Theory}
\label{sec:theory}
Consider a zero mean wide sense stationary, vector valued and discrete
time stochastic process $x(t) \in L^n_2(\Omega);\ t \in \Z$, having the matrix
valued covariance sequence
$R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$, which we assume to be
absolutely summable, and spectra $S(\omega)$:

\begin{equation}
  \label{eqn:fourier_pair}
  \begin{aligned}
    R(\tau) &= \frac{1}{2\pi}\int_{-\pi}^\pi S(\omega) e^{j\tau\omega}\d \omega,\\
    S(\omega) &= \sum_{\tau=-\infty}^\infty R(\tau)e^{-j\tau\omega}.
  \end{aligned}
\end{equation}

\begin{definition}[Granger Causality]
  For the WSS series $x(t)$ we say that component $x_j$
  \textit{Granger-Causes} (GC) component $x_i$ (with respect to $x$)
  and write $x_j \gc x_i$ if given Hilbert spaces
  $\mathcal{H}_{t - 1}$, $\mathcal{H}^{-j}_{t - 1}$
  generated by $\{x_i(t - \tau)\}_{(\tau = 1, i = 1)}^{(\infty, n)}$ and
  $\{x_i(t - \tau)\}_{(\tau = 1, i \ne j)}^{(\infty, n)}$ respectively we have

\begin{equation}
  \linEerr{x_i(t)}{\mathcal{H}_{t - 1}} < \linEerr{x_i(t)}{\mathcal{H}^{-j}_{t - 1}},
\end{equation}

where
$\xi[x \ |\ \mathcal{H}] = \E (x - \linE{x}{\mathcal{H}})^2$ and
$\linE{x}{\mathcal{H}} = \text{proj}_{\mathcal{H}}(x)$ denotes
the (unique) projection onto the Hilbert space $\mathcal{H}$.
\end{definition}

This notion captures the idea that the process $x_j$ provides
information about $x_i$ that is not available from elsewhere.  The
caveat ``with respect to $x(t)$'' is important in that GC relations
can change when components are added to or removed from $x(t)$,
e.g. new GC relations can arise if we remove the observations of a
common cause, and existing GC relations can disappear if we observe a
new mediating series.

The notion is closely related to the information theoretic measure of
transfer entropy, indeed, if everything in Gaussian then they are
equivalent \cite{barnett2009granger}.

We continue to review some pertinent facts.

As a consequence of the Wold decomposition theorem \cite{lindquist},
every such sequence has the ``moving average'' $MA(\infty)$ representation

\begin{equation}
\label{eqn:wold}
  x(t) = c(t) + \sum_{\tau = 0}^\infty A(\tau) z(t - \tau),
\end{equation}

where $c(t)$ is a perfectly predictable sequence\footnote{$c(t)$ is
  perfectly predictable if any sample $c(t_0)$ is enough to determine
  $c(t)$ for every $t$.  For example,
  $c(t) = \text{sin}(2\pi t + \Theta);\; \Theta \sim \mathcal{U}[-\pi, \pi]$}, and
$z(t)$ is an uncorrelated sequence having covariance $I$.  We will assume
that $c(t)$ can be effectively estimated and removed and take
$c(t) = 0$.  Furthermore, we assume that the spectra is uniformly
bounded away from $0$:
$\exists \delta:\ \delta I \preceq S(\omega) \preceq \frac{1}{\delta} I\ \forall \omega$ so that the Wold
representation can be inverted to yield the $VAR(\infty)$ form

\begin{equation}
  \label{eqn:ar_representation}
  x(t) = \sum_{\tau = 1}^\infty B(\tau) x(t - \tau) + v(t),
\end{equation}

where $v(t)$ is an uncorrelated sequence with covariance $\Sigma_v$.

\begin{theorem}[Granger Causality Equivalences]
  Let $x(t)$ be a WSS process with absolutely summable covariance
  sequence, and uniformly bounded spectral density.  Denote by
  $\xi_{ij} \defeq \linEerr{x_i(t)}{\mathcal{H}^{-j}}$ and
  $\xi_i \defeq \linEerr{x_i(t)}{\mathcal{H}_t}$.  Then, the following are equivalent:

  \begin{enumerate}
    \item{$x_j \not\gc x_i$}
    \item{$\forall \tau \in \N_+\ B_{ij}(\tau) = 0$}
    \item{$\mathcal{F}_{j \rightarrow i} \defeq \ln\frac{\xi_i}{\xi_{ij}} = 0$}
    \item{$\mathcal{H}_t^{i} \perp \mathcal{H}_{t - 1}^{j}\ |\ \mathcal{H}_{t - 1}^{-j}$}
    \item{\hl{Wold $A(\tau)$ condition}}
    \item{Tie in the 0s of $S(\omega)^{-1}$, though this isn't immediately directed.}
  \end{enumerate}
\end{theorem}

The equivalence $(1) \iff (3)$ is essentially a restatement of the
definition, but is in line with the seminal work of Geweke
\cite{geweke1982}, \cite{geweke1984}.  Statement (4) means that
$(y - \linE{y}{\mathcal{H}_{t - 1}^{-j}}) \perp (z -
\linE{z}{\mathcal{H}_{t - 1}^{-j}})\ \forall y \in \mathcal{H}_t^i,\ z \in
\mathcal{H}_{t - 1}^j$, which is an appealing geometric statement that
there is no direct feedback from $x_j$ to $x_i$ \cite{lindquist}.

\begin{proof}
$(1) \implies (2)$: The projection for $x_i(t)$ onto $\mathcal{H}_{t - 1}$ is (by definition) given by a form similar to equation \ref{eqn:ar_representation}:

\[
  \E|x_i(t) - \linE{x_i(t)}{\mathcal{H}_{t - 1}}|^2 = \E|v_i(t) + \sum_{\tau = 1}^\infty \sum_{k = 1}^n (B_{i, k}(\tau) - \hat{B}_{i, k}(\tau))x_k(t - \tau)|^2.
\]

 Since $v(t)$ in equation \ref{eqn:ar_representation} is temporally uncorrelated, it follows that the optimal projection is given by the model coefficients themselves.  This holds similarly for the projection onto $\mathcal{H}_{t - 1}^{-j}$.  Then by $(1)$ we have

\[
  \E|x_i(t) - \sum_{\tau = 1}^\infty \sum_{k = 1}^n B_{i, k}(\tau)x_k(t - \tau)|^2 = \E|x_i(t) - \sum_{\tau = 1}^\infty\sum_{k \ne j}B_{i, k}(\tau)x_k(t - \tau)|^2
\]

By the uniqueness of the projection we must have $\forall \tau\  B_{i, j}(\tau) = 0$.

$(2) \implies (4)$: In computing $(y - \linE{y}{\mathcal{H}_{t = 1}^{-j}})$ for $y \in \mathcal{H}_t^i$ it is sufficient to consider $y = x_i(t)$ since $\mathcal{H}_{t - 1}^i \subseteq \mathcal{H}_{t - 1}^{-j}$, in which case $(x_i(t) - \linE{x_i(t)}{\mathcal{H}_{t = 1}^{-j}}) = v_i(t)$.  $(4)$ follows since $v_i(t) \perp \mathcal{H}_{t - 1}$ and $\forall z \in \mathcal{H}_{t - 1}^j\ (z - \linE{z}{\mathcal{H}_{t - 1}^{-j}}) \in \mathcal{H}_{t - 1}$

\end{proof}

The following propositions justify various modifications of $x(t)$
applied in practice to massage $x(t)$ into a form amenable to more
standard tools.

\begin{theorem}[General Invariance]
  \hl{Under what conditions is this actually true?}

  Let $\zeta(t)$ be a stationary discrete time stochastic process.  Let
  $F, G$ be (possibly nonlinear and time varying) invertible filtering
  operations and $f(t)$, $g(t)$ be perfectly predictable and such that
  $x_j(t) \defeq F(\zeta_j - f)(t), x_i(t) \defeq G(\zeta_i - g)(t)$ are W.S.S.  Then,

\begin{equation}
    x_j \gc x_i \iff \zeta_j \te \zeta_i
  \end{equation}
\end{theorem}

\begin{corollary}[Invariance Under Invertible and Deterministic Modifications]
  Let $F(z), G(z)$ describe univariate and invertible
  linear-time-invariant filters.  And, let $f(t), g(t)$ be perfectly
  predictable processes.  Then,

  \begin{equation}
    x_j \gc x_i \iff F(x_j - f)(t) \gc G(x_i - g)(t)
  \end{equation}
\end{corollary}

\subsection{Pairwise Granger Causality}
Recall that Granger-causality in general must be understood with
respect to a particular universe of observations.  If $x_j \gc x_i$
with respect to $x_{-k}$, it may not hold with respect to $x$.  That
is, $x_k$ may be a common ancestor that when observed completely
explains the connection from $x_j$ to $x_i$.  In particular, we will
study \textit{pairwise} Granger-causality, where $x_i \pwgc x_j$ if
$x_i$ Granger-causes $x_j$ with respect only to $(x_i, x_j)$.

We establish first some notation.  Remaining in the context of a
collection of $n$ WSS processes
$x(t) = \big(x_1(t), \ldots, x_n(t)\big)$, we define the Granger-causality
graph $\gcg$ to be the directed graph formed on $n$ vertices where an
edge $(j, i) \in \gcg$ if and only if $x_j \gc x_i$.  We refer to the
set of parents of $x_i$ in $\gcg$ by $\pa{i}$, that is,
$j \in \pa{i} \iff x_j \gc x_i$.  The set of \textit{ancestors} of
$x_i$ in $\gcg$ will be denoted $\anc{i}$, that is, $j \in \anc{i}$ if
and only if there is a directed path in $\gcg$ from $j$ to $i$.  Even
though the definitions of Granger-causality naturally suggests that
each node in the graph should have a self-loop (unless the node is
simply independent noise) we will not include nodes in their own
parent sets: $i \not\in \pa{i} \subseteq \anc{i}$.  We note that $\gcg$ need not be a DAG.

The edges of the graph can be given a general notion of ``weight'' by
associating an edge $(j, i)$ with an LTI filter
$\{B_{ij}(\tau)\}_{\tau = 1}^\infty$ from the $\VAR$ representation of equation
\ref{eqn:ar_representation}.

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
% I'm actually not sure this is true, I think that if there is a common ancestor it is
% ambiguous.  This may be true for transfer entropy, but for Granger-causality even if
% the information arriving at i is later than to j, it may be less corrupt and hence
% still facilitate better predictions about k's future.

Finally, the \textit{propagation delay} from $j$ to $i$, denoted $d(j, i)$, will be defined as

\[
  d(j, i) = \text{min}\big\{\sum_{k = 0}^{K - 1} d(p_k, p_{k + 1}})\ |\ i = p_0 \rightarrow p_1 \rightarrow \cdots \rightarrow p_K = j \text{ is a path in } \gcg \big\}
\]

and $d(j, i) \defeq \text{min}\{B_{i, j}(\tau) \ne 0\}$ if $j \in \pa{i}$.

\begin{lemma}
  \label{lem:pwgc_anc}
  $x_j \pwgc x_i$ if and only if at least one of the following hold

  \begin{enumerate}
    \item{$j \in \anc{i}$}
    \item{$\anc{i}\cap\anc{j} \ne \emptyset$.} %and $d(k, j) < d(k, i)$}
  \end{enumerate}
\end{lemma}

% \begin{remark}
%   This lemma formalizes the intuition of \textit{information flow} in $\gcg$.  We have $x_j \pwgc x_i$ whenever $i$ is descended from $j$, or if they have a common ancestor (call it $k$) and ``information'' from $k$ reaches $j$ before it reaches $i$.
% \end{remark}

\begin{proof}
\end{proof}

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\begin{theorem}[Tree Recovery]
  \label{thm:tree_recovery}
  If the Granger-Causality graph $\gcg$ for $x(t)$ is a
  directed forest, then $\gcg$ can be inferred from pairwise
  causality tests alone.
\end{theorem}

\section{Structure Learning}
\subsection{Local Search Heuristics}

\subsection{Edge-Wise Grouped LASSO}
Minimize the sparsity promoting regularized least squares problem and choose the hyper-parameters against the Akaike information criteria.

\begin{equation}
  L(\lambda, p) \defeq \underset{B}{\text{minimize}}\ \frac{1}{T}\sum_{t = 1}^T||x(t) - \sum_{\tau = 1}^pB(\tau)x(t - \tau)||_2^2 + \lambda \sum_{i, j}\big[\alpha||B_{i, j}||_2 + (1 - \alpha)||B_{i, j}||_1\big]
\end{equation}

\begin{equation}
  \underset{\lambda, p}{\text{minimize}}\ L(\lambda, p) + \mathsf{AIC}(B^{(\lambda, p)})
\end{equation}

\subsection{Bayesian Posterior Thresholding}
Consider the Bayesian model

\begin{equation}
  \begin{aligned}
    (x(t)\ |\ B, \sigma_v^2, \{x(t - \tau)\}_{\tau = 1}^p) &\sim \mathcal{N}(\sum_{\tau = 1}^pB(\tau)x(t - \tau), \sigma_v^2)\\
    (B_{ij}(\tau)\ |\ G_{ij}(\tau)) &\sim G_{ij}(\tau)\mathcal{N}(0, \sigma_\beta^2) + (1 - G_{ij}(\tau))\delta_0(B_{ij}(\tau))\\
    (G_{ij}(\tau)) &\sim \mathsf{BER}(p_{ij}(\tau))\\
    \sigma_v^2 &\sim \Gamma(a_v, b_v)\\
    \sigma_\beta^2 &\sim \Gamma(a_\beta, b_\beta)
  \end{aligned}
\end{equation}

Similar models have been studied by various authors in the context of the linear regression model \hl{[(cite them)]} where it is referred to as \textit{stochastic variable selection}, or referred to as a model for Bayesian variable selection.

Since we have in mind applications where $n$ is large, sampling posterior probabilities can be prohibitively burdensome, so we can instead fit the mean-field variational approximation $p(B, G, \sigma\ |\ X) \approx q_Gq_Bq_\sigma$ and subsequently apply a thresholding operation to the posterior edge inclusion probabilities under $q_G$.

\subsection{Pairwise Minimum Error Spanning Trees}
Inspired by theorem \ref{thm:tree_recovery}, we propose the following
structure learning heuristic.

%% This algorithm should perform coordinate descent or jointly refit all the trees.

\begin{enumerate}
  \item{Set $k = 0$ and $x^0(t) = x(t)$}
  \item{Compute all pairwise causality measures $\Xi \defeq \big[\ln \frac{\xi_i}{\xi_{ij}} \big]_{i, j}$}
  \item{Find the maximum spanning arborescence\footnote{An ``arborescence'' is a french word for a tree diagram, and refers to a \textit{directed} tree in graph theory} $\mathcal{T}_k$ of a graph having edge weights $\Xi_{ij}$.}
  \item{Fit a vector LTI filter $F_k$ having graph defined by $\mathcal{T}_k$ to $x^{k - 1}(t)$ and set $x^k(t) = x^{k - 1}(t) - \hat{x}^{k - 1}(t)$}
  \item{Repeat over $k$ until a stopping criteria is met}
\end{enumerate}

\section{Application}
\section{Conclusion}

\printbibliography
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: