### 2019-03-15 paper.tex
Work on paper -- ancestor expansion.

\documentclass[12pt]{article}

%% bibliography stuff -- this needs come before the preamble inclusion
\usepackage[backend=bibtex,sorting=none]{biblatex}
\bibliography{\string~/Documents/academics/global_academics/global_bib.bib}
\usepackage{hyperref}

\def\gc{\overset{\text{GC}}{\rightarrow}}  % Granger causality arrow
\def\pwgc{\overset{\text{PW}}{\rightarrow}}  % Granger causality arrow
\def\te{\overset{\mathcal{T}}{\rightarrow}}  % Transfer entropy arrow
\def\gcg{\mathcal{G}}  % Granger-causality graph
\def\VAR{\mathsf{VAR}}  % VAR(p) model
\def\B{\mathsf{B}}  % Filter B
\def\wtB{\widetilde{\B}}  % General filter B
\def\A{\mathsf{A}}  % Filter A

\newcommand{\linE}[2]{\hat{\E}[#1\ |\ #2]}  % Linear projection
\newcommand{\linEerr}[2]{\xi[#1\ |\ #2]}  % Error of linear projection
\newcommand{\pa}[1]{pa(#1)}  % Parents of a node
\newcommand{\anc}[1]{\mathcal{A}(#1)}  % Ancestors of a node
\newcommand{\ancn}[2]{\mathcal{A}_{#1}(#2)}  % nth ancestors of a node
\newcommand{\gpn}[2]{gp_{#1}(#2)}  # nth generation grandparents
\newcommand{\gcgpath}[4]{#1 \rightarrow #2_1 \rightarrow \cdots \rightarrow #2_{#3} \rightarrow #4}

\input{\string~/Documents/academics/global_academics/latex_preamble}

\graphicspath{{./figures/}}

\title{Structure Learning for $VAR(p)$ Models}
\author{R. J. Kinnear, R. R. Mazumdar}

\begin{document}
\maketitle
\abstract{We study Granger Causality and propose a structure learning
  heuristic for uncovering a parsimonious representation of large
  $\mathsf{VAR}(p)$ models.}

\section{Introduction and Review}
\label{sec:introduction}
Consider a collection of stochastic processes producing observations
at discrete time intervals.  Are the underlying processes dependent?
Can we quantify any of the underlying relationships?  Can the arrow of
time help us to distinguish a directionality or flow of dependence
among our observed series?  In this paper we contribute to the
understanding of the notion of Granger-Causality
\cite{granger1969investigating} as a tool for answering these questions.

Though the notion of causality is a philosophically slippery concept,
it is fundamental to the way we understand the world and to the
progress of science in general.  Indeed, without faith in the
consistency of causal interactions the results of experimental science
could not be generalized or applied in any meaningful way.  In the
case of Granger-Causality, we state that if an event '$A$' provides us
with unique (that is, not available anywhere else) information about a
later event $B$, then $A$ must have a causal impact on $B$.  As
opposed to the notion of Causation promoted by Pearl
\cite{pearl2000art}, this is an entirely model-free notion of cause,
and instead leverages the intuition that a cause must precede it's
effect.

In practice, Granger's notion of causation it not a convincing test
for \textit{true} causation, since our statements about causation are
highly dependent upon the data that we are able to observe.  We prefer
instead to interpret Granger causality as a means of uncovering a flow
of ``information'' or ``energy'' through some underlying graph of
interactions.  Though this graph cannot be observed directly, we will
infer it's presence as a latent structure among our observed time
series data.

Finding the ``best'' graph structure consistent with observed data is
generally an extremely challenging problem, though the comparison of
quality between different structures, and hence the notion of
``best'', needs to be quantified.  In applications where we are
interested merely in minimizing the mean squared error of a linear
one-step-ahead predictor, then we will naturally desire an entirely
dense graph of connections, since each edge can only serve to reduce
estimation error.  However, since the number of edges scales
quadratically in $n$, it is imperative to infer a sparse causality
graph for large systems, both to avoid overfitting observed data, as
well as to aid the interpretability of the results.

In \cite{bach2004learning} the authors apply a local search heuristic
to an AIC penalized approximation of the likelihood where at each
iteration an edge is either added, removed, or reversed.  This is a
common approach to combinatorial optimization due to it's simplicity,
but is liable to get stuck in shallow local minima.

A second and wildly successful heuristic is the LASSO regularizer
\cite{tibshirani1996regression}, which can be understood as a natural
convex relaxation to penalizing the count of the non-zero edges.  The
LASSO enjoys strong theoretical guarantees \cite{wainwright2009sharp},
even in the case of time series data \cite{basu2015}
\cite{wong2016lasso}.

From a Bayesian perspective, spike-and-slab priors can be used to
estimate the probability of an edge's inclusion, after which a graph
can be chosen heuristically by thresholding or searching for the MAP
graph structure.

Our contribution is to provide an alternative heuristic based on
minimum spanning trees, inspired by our result in section
\ref{sec:theory} that the true causality graph can be recovered from
pairwise information alone in the case that the underlying graph is a
directed forest.  We compare this algorithm to a grouped LASSO
\cite{yuan2006model}, local search, and a Bayesian posterior
probability thresholding scheme.

%% -Can we also formulate an SDP relaxation?
%% -Fit trees jointly or via coordinate descent?

\section{Theory}
\label{sec:theory}
Consider a zero mean wide sense stationary, vector valued and discrete
time stochastic process $x(t) \in L^n_2(\Omega);\ t \in \Z$, having the matrix
valued covariance sequence
$R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$, which we assume to be
absolutely summable, and spectra $S(\omega)$:

\begin{equation}
  \label{eqn:fourier_pair}
  \begin{aligned}
    R(\tau) &= \frac{1}{2\pi}\int_{-\pi}^\pi S(\omega) e^{j\tau\omega}\d \omega,\\
    S(\omega) &= \sum_{\tau=-\infty}^\infty R(\tau)e^{-j\tau\omega}.
  \end{aligned}
\end{equation}

\begin{definition}[Granger Causality]
  For the WSS series $x(t)$ we say that component $x_j$
  \textit{Granger-Causes} (GC) component $x_i$ (with respect to $x$)
  and write $x_j \gc x_i$ if given Hilbert spaces
  $\mathcal{H}_{t - 1}$, $\mathcal{H}^{-j}_{t - 1}$
  generated by $\{x_i(t - \tau)\}_{(\tau = 1, i = 1)}^{(\infty, n)}$ and
  $\{x_i(t - \tau)\}_{(\tau = 1, i \ne j)}^{(\infty, n)}$ respectively we have

\begin{equation}
  \linEerr{x_i(t)}{\mathcal{H}_{t - 1}} < \linEerr{x_i(t)}{\mathcal{H}^{-j}_{t - 1}},
\end{equation}

where
$\xi[x \ |\ \mathcal{H}] = \E (x - \linE{x}{\mathcal{H}})^2$ and
$\linE{x}{\mathcal{H}} = \text{proj}_{\mathcal{H}}(x)$ denotes
the (unique) projection onto the Hilbert space $\mathcal{H}$.
\end{definition}

This notion captures the idea that the process $x_j$ provides
information about $x_i$ that is not available from elsewhere.  The
caveat ``with respect to $x(t)$'' is important in that GC relations
can change when components are added to or removed from $x(t)$,
e.g. new GC relations can arise if we remove the observations of a
common cause, and existing GC relations can disappear if we observe a
new mediating series.

The notion is closely related to the information theoretic measure of
transfer entropy, indeed, if everything in Gaussian then they are
equivalent \cite{barnett2009granger}.

We continue to review some pertinent facts.

As a consequence of the Wold decomposition theorem \cite{lindquist},
every such sequence has the ``moving average'' $MA(\infty)$ representation

\begin{equation}
\label{eqn:wold}
  x(t) = c(t) + \sum_{\tau = 0}^\infty A(\tau) z(t - \tau),
\end{equation}

where $c(t)$ is a perfectly predictable sequence\footnote{$c(t)$ is
  perfectly predictable if any sample $c(t_0)$ is enough to determine
  $c(t)$ for every $t$.  For example,
  $c(t) = \text{sin}(2\pi t + \Theta);\; \Theta \sim \mathcal{U}[-\pi, \pi]$}, and
$z(t)$ is an uncorrelated sequence having covariance $I$.  We will assume
that $c(t)$ can be effectively estimated and removed and take
$c(t) = 0$.  Furthermore, we assume that the spectra is uniformly
bounded away from $0$:
$\exists \delta:\ \delta I \preceq S(\omega) \preceq \frac{1}{\delta} I\ \forall \omega$ so that the Wold
representation can be inverted to yield the $VAR(\infty)$ form

\begin{equation}
  \label{eqn:ar_representation}
  x(t) = \sum_{\tau = 1}^\infty B(\tau) x(t - \tau) + v(t),
\end{equation}

where $v(t)$ is an uncorrelated sequence with covariance $\Sigma_v$.

\begin{theorem}[Granger Causality Equivalences]
  \label{thm:gcg}
  Let $x(t)$ be a WSS process with absolutely summable covariance
  sequence, and uniformly bounded spectral density.  Denote by
  $\xi_{ij} \defeq \linEerr{x_i(t)}{\mathcal{H}^{-j}}$ and
  $\xi_i \defeq \linEerr{x_i(t)}{\mathcal{H}_t}$.  Then, the following are equivalent:

  \begin{enumerate}
    \item{$x_j \not\gc x_i$}
    \item{$\forall \tau \in \N_+\ B_{ij}(\tau) = 0$}
    \item{$\mathcal{F}_{j \rightarrow i} \defeq \ln\frac{\xi_i}{\xi_{ij}} = 0$}
    \item{$\mathcal{H}_t^{i} \perp \mathcal{H}_{t - 1}^{j}\ |\ \mathcal{H}_{t - 1}^{-j}$}
    \item{\hl{Wold $A(\tau)$ condition}}
    \item{Tie in the 0s of $S(\omega)^{-1}$, though this isn't immediately directed.}
  \end{enumerate}
\end{theorem}

The equivalence $(1) \iff (3)$ is essentially a restatement of the
definition, but is in line with the seminal work of Geweke
\cite{geweke1982measurement}, \cite{geweke1984}.  Statement (4) means that
$(y - \linE{y}{\mathcal{H}_{t - 1}^{-j}}) \perp (z -
\linE{z}{\mathcal{H}_{t - 1}^{-j}})\ \forall y \in \mathcal{H}_t^i,\ z \in
\mathcal{H}_{t - 1}^j$, which is an appealing geometric statement that
there is no direct feedback from $x_j$ to $x_i$ \cite{lindquist}.

\begin{proof}
$(1) \implies (2)$: The projection for $x_i(t)$ onto $\mathcal{H}_{t - 1}$ is (by definition) given by a form similar to equation \ref{eqn:ar_representation}:

\[
  \E|x_i(t) - \linE{x_i(t)}{\mathcal{H}_{t - 1}}|^2 = \E|v_i(t) + \sum_{\tau = 1}^\infty \sum_{k = 1}^n (B_{i, k}(\tau) - \hat{B}_{i, k}(\tau))x_k(t - \tau)|^2.
\]

 Since $v(t)$ in equation \ref{eqn:ar_representation} is temporally uncorrelated, it follows that the optimal projection is given by the model coefficients themselves.  This holds similarly for the projection onto $\mathcal{H}_{t - 1}^{-j}$.  Then by $(1)$ we have

\[
  \E|x_i(t) - \sum_{\tau = 1}^\infty \sum_{k = 1}^n B_{i, k}(\tau)x_k(t - \tau)|^2 = \E|x_i(t) - \sum_{\tau = 1}^\infty\sum_{k \ne j}B_{i, k}(\tau)x_k(t - \tau)|^2
\]

By the uniqueness of the projection we must have $\forall \tau\  B_{i, j}(\tau) = 0$.

$(2) \implies (4)$: In computing $(y - \linE{y}{\mathcal{H}_{t = 1}^{-j}})$ for $y \in \mathcal{H}_t^i$ it is sufficient to consider $y = x_i(t)$ since $\mathcal{H}_{t - 1}^i \subseteq \mathcal{H}_{t - 1}^{-j}$, in which case $(x_i(t) - \linE{x_i(t)}{\mathcal{H}_{t = 1}^{-j}}) = v_i(t)$.  $(4)$ follows since $v_i(t) \perp \mathcal{H}_{t - 1}$ and $\forall z \in \mathcal{H}_{t - 1}^j\ (z - \linE{z}{\mathcal{H}_{t - 1}^{-j}}) \in \mathcal{H}_{t - 1}$

\end{proof}

The following propositions justify various modifications of $x(t)$
applied in practice to massage $x(t)$ into a form amenable to more
standard tools.

\begin{theorem}[General Invariance]
  \hl{Under what conditions is this actually true?}

  Let $\zeta(t)$ be a stationary discrete time stochastic process.  Let
  $F, G$ be (possibly nonlinear and time varying) invertible filtering
  operations and $f(t)$, $g(t)$ be perfectly predictable and such that
  $x_j(t) \defeq F(\zeta_j - f)(t), x_i(t) \defeq G(\zeta_i - g)(t)$ are W.S.S.  Then,

\begin{equation}
    x_j \gc x_i \iff \zeta_j \te \zeta_i
  \end{equation}
\end{theorem}

\begin{corollary}[Invariance Under Invertible and Deterministic Modifications]
  Let $F(z), G(z)$ describe univariate and invertible
  linear-time-invariant filters.  And, let $f(t), g(t)$ be perfectly
  predictable processes.  Then,

  \begin{equation}
    x_j \gc x_i \iff F(x_j - f)(t) \gc G(x_i - g)(t)
  \end{equation}
\end{corollary}

\subsection{Pairwise Granger Causality}
Recall that Granger-causality in general must be understood with
respect to a particular universe of observations.  If $x_j \gc x_i$
with respect to $x_{-k}$, it may not hold with respect to $x$.  For
example, $x_k$ may be a common ancestor which when observed completely
explains the connection from $x_j$ to $x_i$.  In particular, we will
study \textit{pairwise} Granger-causality.

\begin{definition}[Pairwise Granger-causality]
  We will say that $x_j$ pairwise Granger-causes $x_i$ and write
  $x_j \pwgc x_i$ if $x_i$ Granger-causes $x_j$ with respect only to
  $(x_i, x_j)$.
\end{definition}

We first need to establish some graph theoretic notation and
terminology, collected formally in a definition for the reader's
convenient reference.

\begin{definition}[Causality graph, Parents, etc...]

Remaining in the context of a collection of $n$ WSS processes
$x(t) = \big(x_1(t), \ldots, x_n(t)\big)$, we define the Granger-causality
graph $\gcg$ to be the directed graph formed on $n$ vertices where an
edge $(j, i) \in \gcg$ if and only if $x_j \gc x_i$.

We refer to the set of \textit{parents} of $x_i$ in $\gcg$ by
$\pa{i}$, that is, $j \in \pa{i} \iff x_j \gc x_i$.

Even though the definitions of Granger-causality naturally suggests
that each node in the graph should have a self-loop (unless the node
is simply independent noise) we will not include nodes in their own
parent sets: $i \not\in \pa{i} \subseteq \anc{i}$.  However, we do not
immediately assume that $\gcg$ is a DAG, and so we will allow for $i$
to be it's own \textit{ancestor}.

The set of level $\ell$ \textit{grandparents} of node $i$, denoted
$\gpn{\ell}{i}$, is the set such that $j \in \gpn{\ell}{i}$ if and only if
there is a \textit{directed and cycle-free path} of length $\ell$ in
$\gcg$ from $j$ to $i$ \textit{or} if $j = i$ and there is a cycle of
length $n$ starting at $i$ and ending back at $i$, that is, $i$ can be
it's own grandparent.  Clearly, $\pa{i} = \gpn{1}{i}$.

Finally, the set of \textit{level $\ell$ ancestors} of
$i$: $\ancn{\ell}{i} = \bigcup_{\lambda \le \ell}\gpn{\lambda}{i}$ is the set such that
$j \in \ancn{\ell}{i}$ if and only if there is a directed and cycle-free
path of length $\ell$ \textit{or less} in $\gcg$ from $j$ to $i$
\textit{or} if $j = i$ and there is a directed cycle of length $\ell$ or
less from $i$ back to $i$ in $\gcg$.  The set of \textit{all
  ancestors} of $i$ (e.g. $\ancn{n}{i}$) is denoted simply $\anc{i}$.
\end{definition}

The edges of the Granger-causality graph $\gcg$ can be given a general
notion of ``weight'' by associating an edge $(j, i)$ with a
\textit{strictly causal} LTI filter
$\B_{ij}(z) = \sum_{\tau = 1}^{\infty} B_{ij}(\tau)z^{-\tau}$ from the
$\VAR$ representation of equation \ref{eqn:ar_representation}.  We
will represent the action (convolution) of this filter compactly as

\begin{equation}
  \sum_{\tau = 1}^\infty B_{ij}(\tau)x_j(t - \tau) \defeq \B_{ij}(z)x_j(t).
\end{equation}

% What if we take this as the definition of PWGC?  then, in a simply causal graph
% the "normal" definition becomes a theorem.  Or, if lem:pwgc_anc is true then
% maybe that is good enough?

From the $\VAR$ representation of $x(t)$ there is clearly a tight relationship between each node and it's parent nodes, indeed

\begin{equation}
  \label{eqn:parent_expanion}
  x_i(t) = v_i(t) + \B_{ii}(z)x_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t).
\end{equation}

% When is this guaranteed w.r.t. the full system?
Furthermore, if the filter $\B_{ii}$ is invertible, we can write

\begin{equation*}
  \begin{aligned}
    (1 - \B_{ii}(z))x_i(t) &= v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\\
    \implies x_i(t) &= (1 - \B_{ii}(z))^{-1}\big[v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\big]\\
    &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t)
  \end{aligned}
\end{equation*}

In general, we will write $\wtB$ for arbitrary filters maintaining the
convention that $\wtB_{ii}$ is a filter that modifies the future of
$x_i$ from it's past and $\wtB_{ij}$ a filter modifying $x_i$ from the
past of $x_j$.  We will use the convention that the filter
$\widetilde{\B}$ will represent simply ``a filter'' and is not
necessarily the same as the filters serving as edge weights in $\gcg$
(indeed, $\wtB$ will often be $0$), and moreover, we will allow for
$\widetilde{\B}$ to potentially change from place to place and even
from line to line in the same sequence of calculations as it's exact
properties or specification are not important.

We continue to develop the relationship between $x_i$ and it's family tree.

\begin{proposition}
  \label{prop:parent_expanding}
  The component $x_i(t)$ of $x(t)$ can be represented as a linear
  combination of a linear filtering operation acting on the driving
  noise $v_i$ of $x_i$ as well as the parents of $x_i$:

  \begin{equation}
    \label{eqn:parent_expansion}
    x_i(t) = \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t).
  \end{equation}

  Moreover, $x_i$ can be expanded in terms of it's ancestors and
  driving noise only:

  \begin{equation}
    x_i(t) = \wtB_{ii}(z)v_i(t) + \sum_{k \in \anc{i} \setminus\{i\}}\wtB_{ik}(z)v_k(t).
  \end{equation}
\end{proposition}
\begin{proof}
  The formula \ref{eqn:parent_expansion} was established in the
  preceding paragraph and serves as a starting point for the induction
  hypothesis:

  \begin{equation}
    \label{eqn:induction}
    x_i(t) = \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell - 1}{i}\setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t).
  \end{equation}

  The base case follows by expanding \ref{eqn:parent_expansion}, but we skip the calculations since they are nearly identical to the following induction step:

  \begin{align*}
    x_i(t) &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell - 1}{i}\setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t)\\
           &\overset{(a)}{=} \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell - 1}{i}\setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)\big[\wtB_{kk}v_k(t)\\
    &\ \ \ \  + \wtB_{ki}(z)x_i(t) + \sum_{h \in \pa{k}\setminus \{i\}}\wtB_{kh}(z)x_h(t)\big]\\
    &\overset{(b)}{=} \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell + 1}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t).
  \end{align*}

  Where set $(a)$ follows by expanding each $x_k$ with equation \ref{eqn:parent_expansion} and $(b)$ requires the inversion of $\big(1 - \sum_{k \in \gpn{\ell}{i}\setminus\{i\}}\wtB_{ik}(z)\wtB_{ki}(z)\big)$ as well as using the fact that $\ancn{\ell}{i} = \ancn{\ell - 1}{i}\cup\gpn{\ell}{i}$.  We here remind the reader of our earlier established convention that $\wtB$ simply represents the existence of a filter (possibly 0) and can change from place to place.

The induction terminates after $n$ steps since in a graph with $n$ nodes $\gpn{n + 1}{i} = \emptyset$

% Where step $(a)$ requires the inversion of $1 - \sum_{k \in \pa{i}}\wtB_{ik}(z)\wtB_{ki}(z)$.

%   \begin{align*}
%     x_i(t) &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t)\\
%            &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)\big[\wtB_{kk}v_{k}(t) + \wtB_{ki}(z)x_i(t) + \sum_{h \in \pa{k}\setminus \{i\}}\wtB_{kh}(z)x_h(t) \big]\\
%            &\overset{(a)}{=} \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)v_{k}(t) + \sum_{k \in \gpn{2}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t)
%   \end{align*}

% Where step $(a)$ requires the inversion of $1 - \sum_{k \in \pa{i}}\wtB_{ik}(z)\wtB_{ki}(z)$.

\end{proof}

\begin{proposition}
  \label{prop:ancestor_properties}
  In a general Granger-causality graph $\gcg$, if $j \pwgc i$ then at
  least one of the following must hold

  \begin{enumerate}
    \item{$j \in \anc{i}$}
    \item{$\anc{i} \cap \anc{j} \ne \emptyset$}
  \end{enumerate}
\end{proposition}
\begin{proof}
\end{proof}

It seems intuitive that a converse of \ref{prop:ancestor_properties}
would hold, e.g. $j \in \anc{i} \implies j \pwgc i$.  Unfortunately,
this is not the case in general, as different paths through $\gcg$ can
lead to cancellation.

\begin{example}
  Firstly, on $n = 4$ nodes, ``diamond'' shapes can lead to cancellation on paths of length 2:

\begin{equation*}
  x(t) =
  \left[
    \begin{array}{cccc}
      0 & 0 & 0 & 0\\
      a & 0 & 0 & 0\\
      -a & 0 & 0 & 0\\
      0 & 1 & 1 & 0\\
    \end{array}
  \right] x(t - 1) + v(t),
\end{equation*}

with $\E v(t) = 0,\ \E v(t)v(t - \tau)^\T = \delta_\tau I$.

The graph associated to this system is not strongly causal since there are two paths between $1$ and $4$, namely: $1 \rightarrow 2 \rightarrow 4$ and $1 \rightarrow 3 \rightarrow 4$.  However, by directly calculating

\begin{align*}
  x_4(t) &= x_2(t - 1) + x_3(t - 1) + v_4(t)\\
         &= ax_1(t - 2) + av_2(t - 1) - ax_1(t - 2) -av_3(t - 1) + v_4(t)\\
         &= a(v_2(t - 1) - v_3(t - 1)) + v_4(t),
\end{align*}

we see that, since $v(t)$ is spatially and temporally uncorrelated, $1 \not\pwgc 4$.
\end{example}

\begin{example}
  A second example on $n = 3$ nodes is also worth examining, in this
  cancellation is a result of differing time lags:

\begin{equation*}
  x(t) =
  \left[
    \begin{array}{ccc}
      0 & 0 & 0\\
      -a & 0 & 0\\
      0 & 1 & 0\\
    \end{array}
  \right] x(t - 1) +
  \left[
    \begin{array}{ccc}
      0 & 0 & 0\\
      0 & 0 & 0\\
      a & 0 & 0\\
    \end{array}
  \right] x(t - 2) + v(t)
\end{equation*}

Then

\begin{align*}
  x_2(t) &= v_2(t) - ax_1(t - 1)\\
  x_3(t) &= v_3(t) + x_2(t - 1) + ax_1(t - 2)\\
  \implies x_3(t) &= v_2(t - 1) + v_3(t),
\end{align*}

and again the graph is not strongly causal and $1 \not\pwgc 3$.
\end{example}

The previous examples seem rather pathological since we have simply
constructed a case where different paths in the graph cancel exactly.
It may be possible to rid ourselves of these pathologies for instance
by defining a notion of ``robust'' pairwise causality where
``$j \pwgc _R i$'' if $j \pwgc i$ for any ``small'' perturbation of
the system matrix.  However, we do not believe such a modification of
our theory can have useful practical implications since the difference
between paths that ``nearly cancel'' rather than cancel exactly would
in practice require inordinate amounts of data to resolve.

Alternatively, we could restrict the coefficients of the system
matrix, e.g. by requiring that $B_{ij}(\tau) \ge 0$.  Instead, we will
focus on the defining feature of time series networks, the topology of
$\gcg$ and consider conditions thereon.

\begin{definition}[Strongly Causal]
  We will say that a Granger-causality graph $\gcg$ is \textit{strongly causal}
if $\gcg$ is a DAG and there is at most 1 directed path between any two nodes.
\end{definition}

\begin{proposition}
  \label{prop:sc_graph_common_anc}
  In a strongly causal graph if $j \in \anc{i}$ and $k \in \anc{i} \cap \anc{j}$ then the unique path from $k$ to $i$ contains $j$.
\end{proposition}
\begin{proof}
  Suppose that there is a path from $k$ to $i$ which does not contain $j$.  In this case, there are multiple paths from $k$ to $i$ (one of which \textit{does} go through $j$ since $j \in \anc{i}$) which contradicts the assumption of strong causality.
\end{proof}

% Is pwgc transitive?

In light of proposition \ref{prop:sc_graph_common_anc}, the following provides a converse to proposition \ref{prop:ancestor_properties}.

\begin{proposition}
  \label{lem:pwgc_anc}
  If $\gcg$ is strongly causal and each self-loop filter
  $1 - \B_{ii}(z)$ is invertible, then $j \in \anc{i} \implies j \pwgc i$.
\end{proposition}
\begin{proof}
  Since strongly causal graphs are acyclic, this allows us to
  recursively expand the parents of $x_i$ in $\gcg$ in the
  $\VAR(\infty)$ representation of $x(t)$ since we will eventually
  terminate at nodes $d$ such that $\pa{d} = \emptyset$.  To this end,
  examining $x_i$ in equation \ref{eqn:ar_representation}

  \begin{align*}
    x_i(t) &= v_i(t) + \sum_{\tau = 1}^\infty \big[ B_{ii}(\tau)x_i(t - \tau) + \sum_{k \in \pa{i}} B_{ik}(\tau)x_k(t - \tau)\big]\\
    \implies x_i(t) &= (1 - \B_{ii}(z))^{-1} \big[v_i(t) + \sum_{k \in \pa{i}} B_{ik}(\tau)x_k(t - \tau) \big]\\
    &= \sum_{\tau = 1}^\infty \big[C_{ii}(\tau)v_i(t - \tau) + \sum_{k \in \pa{i}} C_{ik}(\tau)x_k(t - \tau)\big].
  \end{align*}

  If $\pa{i} = \emptyset$ then we simply have
  $x_i(t) = \sum_{\tau = 1}^\infty C_{ii}(\tau)v_i(t)$ and
  $C_{ii} = A_{ii}$ from equation \ref{eqn:wold}.

  Applying this relation recursively we can write

  \begin{equation*}
    x_i(t) = \sum_{\tau = 1}^\infty\big[C_{ii}(\tau)v_i(t - \tau) + \sum_{k \in \anc{i}}C_{ik}(\tau)v_k(t - \tau) \big].
  \end{equation*}

\end{proof}

That strong conditions need to be placed on the topology of $\gcg$ in
order for lemma \ref{lem:pwgc_anc} to be true can be seen through the
earlier examples.

\begin{lemma}
  If $\gcg$ is strongly causal, then any $\VAR(p)$ system on $\gcg$ is stable if and only if
  $B_{ii}(z)$ is stable for every $i = 1, \ldots, n$.
\end{lemma}
\begin{proof}
\end{proof}

% $B(1) = \left[ \begin{array}{cc} 0 & 1 \\ 1 & 0 \end{array} \right]$

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
% I'm actually not sure this is true, I think that if there is a common ancestor it is
% ambiguous.  This may be true for transfer entropy, but for Granger-causality even if
% the information arriving at i is later than to j, it may be less corrupt and hence
% still facilitate better predictions about k's future.

% Finally, the \textit{propagation delay} from $j$ to $i$, denoted $d(j, i)$, will be defined as

% \[
%   d(j, i) = \text{min}\big\{\sum_{k = 0}^{K - 1} d(p_k, p_{k + 1}})\ |\ i = p_0 \rightarrow p_1 \rightarrow \cdots \rightarrow p_K = j \text{ is a path in } \gcg \big\}
% \]

% and $d(j, i) \defeq \text{min}\{B_{i, j}(\tau) \ne 0\}$ if $j \in \pa{i}$.

\begin{lemma}
  \label{lem:pwgc_anc}
  If $x_j \pwgc x_i$ then one or more of the following hold

  \begin{enumerate}
    \item{$j \in \anc{i}$}
    \item{$\anc{i}\cap\anc{j} \ne \emptyset$.}
  \end{enumerate}
\end{lemma}

% \begin{remark}
%   This lemma formalizes the intuition of \textit{information flow} in $\gcg$.  We have $x_j \pwgc x_i$ whenever $i$ is descended from $j$, or if they have a common ancestor (call it $k$) and ``information'' from $k$ reaches $j$ before it reaches $i$.
% \end{remark}

\begin{proof}
  Is this true if we take \ref{lem:pwgc_anc} as an axiom?
\end{proof}

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

We are interested in understanding under what conditions a causality
graph can be correctly recovered by pairwise testing alone.  It is not
enough that the graph simply be sparse, as the following example shows.

\begin{example}
  If the causality graph $\gcg$ of the vector process $x(t)$ is
  strongly connected\footnote{there is a directed path from each node
    to every other node}, then the causality graph recovered via
  pairwise causality tests is the complete graph.

  This follows immediately from the fact (lemma \ref{lem:pwgc_anc})
  that $j \in \anc{i} \implies j \pwgc i$.
\end{example}

\begin{definition}[Simply Causal]
  If the Granger-Causality graph $\gcg$ for $x(t)$ is as a directed
  acyclic graph such that there is at most one non-overlapping
  directed path between any two nodes (excluding self-loops) then we
  will refer to $\gcg$ as being \textit{simply causal}.
\end{definition}

\begin{example}
  A directed tree (more generally directed forest) is a simply causal
  graph.
\end{example}

\begin{lemma}
  If $\gcg$ is simply causal then $i \pwgc j$ and $j \in \anc{i}$ are
  \textit{alternatives}, that is $i \pwgc j \implies j \notin \anc{i}$.
\end{lemma}
\begin{proof}
  Suppose that $\gcg$ is simply causal and that we have both
  $i \pwgc j$ and $j \in \anc{i}$.  We will show that $i \in \anc{j}$,
  which produces a contradiction (since $\gcg$ being a DAG is part of
  the requirements for being simply causal).

  Since $j \in \anc{i}$ there is a path $\gcgpath{j}{a}{p}{i}$.
  Moreover, since $i \pwgc j$ by lemma \ref{lem:pwgc_anc} there must
  be a common ancestor $k \in \anc{i} \cap \anc{j}$ (otherwise, we would
  have $i \in \anc{j}$, contradicting the acyclic property).  The node
  $k$ has paths to both $j,\ \gcgpath{k}{b}{q}{j}$ and $i,\ \gcgpath{k}{c}{r}{i}$.

\end{proof}

\begin{theorem}[Pairwise Recovery]
  \label{thm:tree_recovery}
  If the Granger-Causality graph $\gcg$ for $x(t)$ is simply causal
  then then $\gcg$ can be inferred from pairwise causality tests
  alone.
\end{theorem}
\begin{proof}

\end{proof}

\begin{example}
  The condition of simple causality is merely a sufficient condition.  For example, the complete directed graph with 2 nodes (e.g. $B(1) = \left[ \begin{array}{cc} 0 & 1 \\ 1 & 0 \end{array} \right]$) contains a loop but is pairwise recoverable.
\end{example}

\section{Structure Learning}
\subsection{Local Search Heuristics}

\subsection{Edge-Wise Grouped LASSO}
Minimize the sparsity promoting regularized least squares problem and choose the hyper-parameters against the Akaike information criteria.

\begin{equation}
  L(\lambda, p) \defeq \underset{B}{\text{minimize}}\ \frac{1}{T}\sum_{t = 1}^T||x(t) - \sum_{\tau = 1}^pB(\tau)x(t - \tau)||_2^2 + \lambda \sum_{i, j}\big[\alpha||B_{i, j}||_2 + (1 - \alpha)||B_{i, j}||_1\big]
\end{equation}

\begin{equation}
  \underset{\lambda, p}{\text{minimize}}\ L(\lambda, p) + \mathsf{AIC}(B^{(\lambda, p)})
\end{equation}

\subsection{Bayesian Posterior Thresholding}
Consider the Bayesian model

\begin{equation}
  \begin{aligned}
    (x(t)\ |\ B, \sigma_v^2, \{x(t - \tau)\}_{\tau = 1}^p) &\sim \mathcal{N}(\sum_{\tau = 1}^pB(\tau)x(t - \tau), \sigma_v^2)\\
    (B_{ij}(\tau)\ |\ G_{ij}(\tau)) &\sim G_{ij}(\tau)\mathcal{N}(0, \sigma_\beta^2) + (1 - G_{ij}(\tau))\delta_0(B_{ij}(\tau))\\
    (G_{ij}(\tau)) &\sim \mathsf{BER}(p_{ij}(\tau))\\
    \sigma_v^2 &\sim \Gamma(a_v, b_v)\\
    \sigma_\beta^2 &\sim \Gamma(a_\beta, b_\beta)
  \end{aligned}
\end{equation}

Similar models have been studied by various authors in the context of the linear regression model \hl{[(cite them)]} where it is referred to as \textit{stochastic variable selection}, or referred to as a model for Bayesian variable selection.

Since we have in mind applications where $n$ is large, sampling posterior probabilities can be prohibitively burdensome, so we can instead fit the mean-field variational approximation $p(B, G, \sigma\ |\ X) \approx q_Gq_Bq_\sigma$ and subsequently apply a thresholding operation to the posterior edge inclusion probabilities under $q_G$.

\subsection{Pairwise Minimum Error Spanning Trees}
Inspired by theorem \ref{thm:tree_recovery}, we propose the following
structure learning heuristic.

%% This algorithm should perform coordinate descent or jointly refit all the trees.

\begin{enumerate}
  \item{Set $k = 0$ and $x^0(t) = x(t)$}
  \item{Compute all pairwise causality measures $\Xi \defeq \big[\ln \frac{\xi_i}{\xi_{ij}} \big]_{i, j}$}
  \item{Find the maximum spanning arborescence\footnote{An ``arborescence'' is a french word for a tree diagram, and refers to a \textit{directed} tree in graph theory} $\mathcal{T}_k$ of a graph having edge weights $\Xi_{ij}$.}
  \item{Fit a vector LTI filter $F_k$ having graph defined by $\mathcal{T}_k$ to $x^{k - 1}(t)$ and set $x^k(t) = x^{k - 1}(t) - \hat{x}^{k - 1}(t)$}
  \item{Repeat over $k$ until a stopping criteria is met}
\end{enumerate}

\section{Application}
\section{Conclusion}

\printbibliography
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: