### 2019-09-29 presentations/mcgillOct2019/pres.tex
Presentation updates.

\documentclass{beamer} %With pauses
% \input{\string~/Documents/academics/global_academics/latex_preamble}
%\documentclass[handout]{beamer} %Without pauses
\usepackage[backend=bibtex]{biblatex}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{cancel}
\usepackage{centernot}
\usefonttheme[onlymath]{serif}

\setbeamertemplate{bibliography item}{}
\setbeamertemplate{bibliography entry title}{}
\setbeamertemplate{bibliography entry location}{}
\setbeamertemplate{bibliography entry note}{}

\addtobeamertemplate{navigation symbols}{}{%
    \usebeamerfont{footline}%
    \usebeamercolor[fg]{footline}%
    \hspace{1em}%
    \insertframenumber/\inserttotalframenumber
}

\bibliography{\string~/Documents/academics/global_academics/global_bib.bib}

\usepackage{wasysym}
\usepackage{tikz}

\usetheme{PaloAlto}

\newtheorem*{defn}{Definition}
\newtheorem*{proposition}{Proposition}

\def\E{\mathbb{E}}  % Expectation
\def\gc{\overset{\text{GC}}{\rightarrow}}  % Granger causality arrow
\def\ngc{\overset{\text{GC}}{\nrightarrow}}  % Negated Granger causality arrow
\def\pwgc{\overset{\text{PW}}{\rightarrow}}  % Pairwise Granger causality arrow
\def\npwgc{\overset{\text{PW}}{\nrightarrow}}  % Negated pairwise Granger causality arrow
\def\te{\overset{\mathcal{T}}{\rightarrow}}  % Transfer entropy arrow
\def\gcg{\mathcal{G}}  % Granger-causality graph
\def\gcge{\mathcal{E}}  % Graph edges
\def\VAR{\mathsf{VAR}}  % VAR(p) model
\def\B{\mathsf{B}}  % Filter B
\def\wtB{\widetilde{\B}}  % General filter B
\def\A{\mathsf{A}}  % Filter A
\def\H{\mathcal{H}}  % Hilbert space
\def\R{\mathbb{R}}  % Reals
\def\N{\mathbb{N}}  % Naturals
\def\T{\mathsf{T}}  % Transpose
\def\Z{\mathbb{Z}}  % Transpose
\def\cl{\mathsf{cl}}  % Closure

\newcommand{\linE}[2]{\hat{\E}[#1\ |\ #2]}  % Linear projection
\newcommand{\linEerr}[2]{\xi[#1\ |\ #2]}  % Error of linear projection
\newcommand{\pa}[1]{pa(#1)}  % Parents of a node
\newcommand{\anc}[1]{\mathcal{A}(#1)}  % Ancestors of a node
\newcommand{\ancn}[2]{\mathcal{A}_{#1}(#2)}  % nth ancestors of a node
\newcommand{\gpn}[2]{gp_{#1}(#2)}  % nth generation grandparents
\newcommand{\wtalpha}[2]{\widetilde{\alpha}(#1, #2)}  % Some notation for lem:pwgc_anc
\newcommand{\dist}[2]{\mathsf{d}(#1, #2)}  % Distance between things
\newcommand{\gcgpath}[2]{#1 \rightarrow \cdots \rightarrow #2}  % A shorter path command

\title{Graph Topological Aspects of Granger Causal Network Learning}
\author{\texorpdfstring{Ryan J. Kinnear\newline\url{Ryan@Kinnear.ca}}{Ryan J. Kinnear}}% \author{Ryan J. Kinnear} \\
  % \small\href{mailto:ryan@kinnear.ca}{ryan@kinnear.ca} \\
  % \small\url{https://github.com/RJTK}\\
  % Under the Supervision of Ravi R. Mazumdar}

\institute[University of Waterloo] % (optional, but mostly needed)
{
  University of Waterloo\\
  Department of Electrical and Computer Engineering
}

\date{McGill University, Oct 4 2019}

\AtBeginSection[]
{
  \begin{frame}<beamer>{Outline}
    \tableofcontents[currentsection,currentsubsection]
  \end{frame}
}

% Let's get started
\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
  % You might wish to add the option [pausesections]
\end{frame}

\section{Introduction}

\begin{frame}{Causality}
  \begin{itemize}
    \item{Do billiard collisions cause motion?}\pause
    \item{Does standing in the street cause cars to stop?}\pause
    \item{Does smoking cause cancer?}\pause
    \item{Does government spending cause an increase in GDP?}\pause
    \item{Granger causality attemps to address the last question's category.}
  \end{itemize}
\end{frame}

\begin{frame}{Causality in Time Series}
  \begin{itemize}
    \item{Consider $x_t, y_t \overset{\text{i.i.d.}}{\sim} \text{BER}(1/2)$}\pause
    \item{If $z_t = x_{t - 1}$ then since cause preceeds effect, intuitively $x$ causes $z$}\pause
    \item{$z'_t = x_{t - 1} \oplus y_{t - 1}$ is more ambiguous.}\pause
    \item{Issue raised by: \fullcite{transfer_entropy_criticism}} 
  \end{itemize}
\end{frame}

\begin{frame}{Causality in Linear Time Series}
  \begin{itemize}
    \item{Granger causality formalizes ``cause proceeds effect'' for WSS processes}\pause
    \item{Interpretation as a ``Flow of Information'' or ``Flow of Energy'' is also reasonable}\pause
    \item{Given WSS processes, $x_1(t), \ldots, x_n(t)$, how can we understand causation $x_i \rightarrow x_j$?}\pause
    \item{\fullcite{granger1969investigating}}
  \end{itemize}
\end{frame}

\begin{frame}{Basic Setup}
  \begin{itemize}
    \item{Basic setup is an $n-$dimensional WSS process $x(t)$}\pause
    \item{Wold representation $x(t) = \sum_{\tau = 0}^\infty A(\tau)v(t - \tau)$}\pause
    \item{$\text{VAR}(\infty)$ representation $x(t) = \sum_{\tau = 1}^\infty B(\tau)x(t - \tau) + v(t)$}\pause
    \item{Require $\E[v(t)v(t - \tau)] = \delta_{\tau}\Sigma_v$, with $\Sigma_v$ Diagonal}\pause
  \end{itemize}

  \begin{defn}
    Linear Projection onto Hilbert space $\H \subseteq L_2(\Omega)$
    \begin{equation*}
      \linE{x}{\H} = \underset{u \in \H}{\text{argmin }} ||x - u||
    \end{equation*}

    Minimum Mean Squared Error
    \begin{equation*}
      \xi[x \ |\ \H] = \E (x - \linE{x}{\H})^2      
    \end{equation*}
  \end{defn}
\end{frame}

\begin{frame}{Basic Setup}
  More notation

  \begin{align*}
    \H_t &= \cl \{\sum_{\tau = 0}^p a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, p \in \N\} \subseteq L_2(\Omega)\\
    % H_t^x &= \{a x(t)\ |\ a \in \R\} \subseteq L_2(\Omega),
  \end{align*}\pause

  \begin{itemize}
    \item{$\H_t^{i}$ Hilbert space spanned by $x_i(t)$ alone}
    \item{$\H_t^{-i}$ Hilbert space spanned by $x(t)$ but \textit{excluding} $x_i(t)$}
  \end{itemize}
\end{frame}

\begin{frame}{Granger Causality}
  \begin{defn}
    For the WSS series $x(t)$ we say that component $x_j$
    \textit{Granger-Causes} component $x_i$ (with respect to $x$)
    and write $x_j \gc x_i$ if given Hilbert spaces $\H_{t - 1}$,
    $\H^{-j}_{t - 1}$ if

    \begin{equation*}
      \linEerr{x_i(t)}{\H_{t - 1}} < \linEerr{x_i(t)}{\H^{-j}_{t - 1}},
    \end{equation*}
  \end{defn}\pause

  \begin{defn}
    We say that $x_j$ pairwise Granger-causes $x_i$ and write
    $x_j \pwgc x_i$ if $x_j$ Granger-causes $x_i$ with respect only to
    $(x_i, x_j)$.
  \end{defn}
\end{frame}

\begin{frame}{Simple Examples}
  \begin{equation*}
    x_i(t) = b_i x(t - 1) + v_i(t); i = 0, 1, 2
  \end{equation*}

  \begin{figure}
    \includegraphics[width=0.75\linewidth]{./figures/var1_example.pdf}
  \end{figure}
\end{frame}

\begin{frame}{Simple Examples}
  \begin{align*}
    x_0(t) &= b_0 x_0(t - 1) + b_{0, 2} x_2(t - 150) + v_0(t)\\
    x_1(t) &= b_1 x_1(t - 1) + b_{1, 2} x_2(t - 150) + v_1(t)\\
    x_2(t) &= b_2 x_2(t - 1) + v_2(t)\\
  \end{align*}

  % TODO: Fix the whitespace consumed by this figure!

  \centering
  \includegraphics[width=0.75\linewidth]{./figures/var2_example.pdf}
\end{frame}

\section{Causality Graphs}
\begin{frame}{Graph Terminology}
  \begin{itemize}
    \item{Graph $\gcg = (V, \gcge)$ with nodes (or vertices) $V$ and edges $\gcge$}\pause
    \item{Directed paths: $j \rightarrow a_0 \rightarrow \cdots \rightarrow a_r \rightarrow i$ for $(a_{k}, a_{k + 1}) \in \gcge$ distinct nodes}\pause
    \item{Cycles: paths with $i = j$}
    \item{Directed Acyclic Graph (DAG): A directed graph without any cycles}\pause
    \item{Parents: $j \in \pa{i}$ iff $(j, i) \in \gcge$}\pause
    \item{Ancestors: $j \in \anc{i}$ iff $j \rightarrow i$ path in $\gcge$}\pause
  \end{itemize}
\end{frame}

\begin{frame}{Granger Causality Graph}
  A Granger causality graph is simply a graph $\gcg = ([n], \gcge)$
  with $(j, i) \in \gcge$ iff $j \gc i$.  $\VAR(1)$ models are
  particularly simple:

  \begin{columns}
    \begin{column}{0.25\textwidth}
      \includegraphics[width=1.5\linewidth]{../../figures/example1.pdf}
    \end{column}

    \begin{column}{0.75\textwidth}
      \begin{equation*}
        x(t) = \left[
          \begin{array}{cccc}
            0 & 0 & 0 & 0\\
            a & 0 & 0 & 0\\
            -a & 0 & 0 & 0\\
            0 & 1 & 1 & 0\\
          \end{array}\right]
        x(t - 1) + v(t)
      \end{equation*}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Application Examples}
  Analysis of connectedness of financial institutions

  TODO: Finance example
  \begin{columns}

  \end{columns}

  \footfullcite{NBERw16223}
\end{frame}

\begin{frame}{Application Examples}
  Analysis of brain functional connectivity

  TODO: Anna examples or something
\end{frame}

\begin{frame}{Application Examples}
  TODO: Gene expression networks
\end{frame}

\section{Recovery Methods}
\begin{frame}{LASSO Recovery Methods}
  \begin{itemize}
    \item{Small systems ($n = 2$ or $n \lessapprox 10$) with adequate data are amenable to least squares and hypothesis testing}\pause
    \item{Large systems ($n \gtrapprox 20$) or when data is limited are addressed via the LASSO (and variants)}\pause
    \item{Standard linear regression case:
        \begin{equation*}
          \underset{\beta \in \R^n}{\text{minimize }} \frac{1}{2T}||y - X\beta||_2^2 + \lambda ||\beta||_1
        \end{equation*}
      }\pause
    \item{$\VAR(p)$ models:
        \begin{equation*}
          \underset{B(\tau) \in \R^{n \times n}}{\text{minimize }} \frac{1}{2T}||x(t) - \sum_{\tau = 1}^p B(\tau)x(t - \tau)||_2^2 + \lambda \sum_{\tau = 1}^p ||B(\tau)||_1
        \end{equation*}
      }
  \end{itemize}
\end{frame}

\begin{frame}{LASSO Recovery Theorems}
  TODO: Static Case
\end{frame}

\begin{frame}{LASSO Recovery Theorems}
  TODO: VAR(p) case
\end{frame}

\begin{frame}{Structure of Granger Causality Graphs}
  \begin{itemize}
    \item{Topology of Granger causality graph is a special feature of time series}\pause
    \item{TODO: Star shaped graph with a self-loop versus a full cycle with graph -- same number of edges, but qualitatively totally different.}
  \end{itemize}
\end{frame}
% \item{Confounder: A node $k$ confounds $i, j$ if $k \in \anc{i} \cap \anc{j}$ if exists a
%     path $\gcgpath{k}{i}$ not containing $j$, and a path
%     $\gcgpath{k}{j}$ not containing $i$.}

\section{Pairwise Causality and ``Causal Flow''}
\begin{frame}{Pairwise Testing}
  \begin{itemize}
    \item{Statistical tests for the case $n = 2$ are simple and well established}\pause
    \item{Can we recover all of $\gcg$ by only testing pairwise relations?}\pause
    \item{Thresholding schemes suffer from confounding}\pause
    \item{``Naive'' pairwise testing is so bad, few papers exist on the topic, or it is used as a simple baseline}\pause
    \item{\fullcite{tam2013gene_pwgc}}
  \end{itemize}
\end{frame}

\begin{frame}{Pairwise Causality}
  \begin{itemize}
    \item{Recall $j \pwgc i$ if $x_j$ Granger causes $x_i$ w.r.t $(x_i, x_j)$}\pause
  \end{itemize}

  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
        \item{Intuition suggests that $j \in \anc{i} \implies j \pwgc i$}\pause
      \end{itemize}

      \includegraphics[width=0.75\linewidth]{../../figures/example1.pdf}\pause
    \end{column}

    \begin{column}{0.5\textwidth}
      \begin{itemize}
        \item{At least $j \gc i \implies j \pwgc i$}\pause
      \end{itemize}
      \includegraphics[width=0.75\linewidth]{../../figures/example2.pdf}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Necessary Conditions for Pairwise Causality}
  First, when $i, j$ are completely separated:

  \begin{proposition}
    Consider distinct nodes $i, j$ in a Granger causality graph
    $\gcg$.  If

    \begin{enumerate}
      \item{$j \not\in \anc{i}$ and $i \not\in \anc{j}$}\pause
      \item{$\anc{i}\cap\anc{j} = \emptyset$}\pause
    \end{enumerate}

    then $\H_t^{(i)} \perp \H_t^{(j)}$: $\forall s, \tau \in \Z_+\ \E[x_i(t - s)x_j(t - \tau)] = 0$.
    \pause  Moreover, this means that $j \npwgc i$ and $\linE{x_j(t)}{\H_t^{(i)}} = 0$.\pause
  \end{proposition}

  Note that $i, j$ may still share descendants.
\end{frame}

\begin{frame}{Necessary Conditions for Pairwise Causality}
  $\H_t^{(i)} \perp \H_t^{(j)}$ is a much stronger conclusion that $j \npwgc i$:
  \begin{proposition}
    Consider distinct nodes $i, j$ in a Granger causality graph $\gcg$.
    If

    \begin{enumerate}
      \item{$j \not\in \anc{i}$}\pause
      \item{$\anc{i}\cap\anc{j} = \emptyset$}\pause
    \end{enumerate}

    then $j \npwgc i$.
  \end{proposition}\pause

  \begin{itemize}
    \item{What about the simple graph $k \rightarrow i \rightarrow j$?}\pause
    \item{Surely $j \npwgc i$, even though $k \in \anc{i} \cap \anc{j}$?}
  \end{itemize}
\end{frame}

\begin{frame}{Necessary Conditions for Pairwise Causality}
  \begin{definition}[Confounder]
    A node $k$ is called a \textit{confounder} of nodes $i, j$ if
    $k \in \anc{i} \cap \anc{j}$ and there exists a path
    $\gcgpath{k}{i}$ not containing $j$, and a path $\gcgpath{k}{j}$
    not containing $i$.  A simple example is furnished by the ``fork''
    graph $i \leftarrow k \rightarrow j$.
  \end{definition}\pause

\begin{proposition}
  \label{prop:ancestor_properties}
  If in a Granger causality graph $\gcg$ where $j \pwgc i$ then
  $j \in \anc{i}$ or $\exists k \in \anc{i} \cap\anc{j}$ which is a
  confounder of $(i, j)$.
\end{proposition}
\end{frame}

\begin{frame}{Sufficient Conditions for Pairwise Causality}
  \begin{itemize}
    \item{Unfortunately, there is no general converse (i.e. $j \in \pa{i} \centernot\implies j \pwgc i$)}\pause
    \item{We seek special (graph-)\textit{topological} conditions when there is a converse}\pause
  \end{itemize}

  \begin{definition}[Strongly Causal]
    \label{def:strongly_causal}
    We will say that a Granger causality graph $\gcg$ is
    \textit{strongly causal} if there is at most one directed path between
    any two nodes.  Strongly Causal Graphs will be referred to as SCGs.
  \end{definition}
\end{frame}

\begin{frame}{Strongly Causal Graph Examples}
  \begin{columns}
    \begin{column}{0.33\linewidth}
      TODO: Simple example\pause
    \end{column}
    \begin{column}{0.33\linewidth}
      \small{Slightly larger example}
      \includegraphics[width=\linewidth]{../../figures/example_scg.pdf}\pause
    \end{column}
    \begin{column}{0.33\linewidth}
      \small{Similar number of edges, but not strongly causal}
      \includegraphics[width=\linewidth]{../../figures/example_dag.pdf}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{(Almost) Strongly Causal Graph Examples}
  \begin{columns}
    \begin{column}{0.5\linewidth}
      \small{E. Coli gene network \tiny{(\fullcite{discovering_graphical_Granger_causality_using_the_truncating_lasso_penalty})}}
      \includegraphics[width=\linewidth]{../../figures/ecoli_regulatory_network.pdf}\pause
    \end{column}
    \begin{column}{0.5\linewidth}
      \small{Saccharomyces cerevisiae gene network \tiny{(\fullcite{learning_genome_scale_regulatory_networks})}}
      \includegraphics[width=\linewidth]{../../figures/huge_gene_network.pdf}\pause
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{A Partial Converse}
  \begin{itemize}
    \item{In strongly causal (and acyclic) graphs, we have the appealing results:}
  \end{itemize}

  \begin{proposition}
    \label{prop:pwgc_anc}
    If $\gcg$ is a strongly causal DAG then $j \in \anc{i} \Rightarrow j \pwgc i$.
  \end{proposition}\pause

  \begin{corollary}
    \label{cor:gc_implies_pwgc}
    If $\gcg$ is a strongly causal DAG then $j \gc i \Rightarrow j \pwgc i$.
  \end{corollary}\pause

  \begin{itemize}
    \item{Unfortunately, effects of confounding more nuanced}\pause
  \end{itemize}
\end{frame}

\begin{frame}{Persistent Systems}
  \begin{itemize}
    \item{If $k \in \anc{i}\cap\anc{j}$ any of the following may occur:}\pause
    \item{
        $(i \npwgc j$, $j \npwgc i)$,
        $(i \pwgc j$, $j \npwgc i)$,
        $(i \npwgc j$, $j \pwgc i)$,
        $(i \pwgc j$, $j \pwgc i)$}\pause
    \item{Intuitively, $i$ may ``forget'' information from $k$ before the same information reaches $i$.}\pause
  \end{itemize}

  A heuristic ``definition'' (to avoid introducing more cumbersome notation):
  \begin{definition}[Persistent]
    We will say that the process $x(t)$ with Granger causality graph
    $\gcg$ is \textit{persistent} if every node has a self-loop,
    i.e. $\B_{ii}(z) \ne 0$.
  \end{definition}

\end{frame}

\section{Pairwise Testing Recovery Methods}
\begin{frame}
\end{frame}

\end{document}